Software Citation Principles

### 2016-03-08 sc-principles/software-citation-principles.tex
more writing in 5

The principles in this section are written fairly concisely, and further discussed later in this document (in \S\ref{sec:examples}). For example, in this section, we are not defining what software should be cited, but how it should be cited.  In the next section, we discuss what how we think such decisions might be made.

@@

\subsection{Discussion of principles}

\paragraph{Derived Software}

@@

\paragraph{what software to cite}

\katznote{put discussion of what software to cite here -- something about what type of software we think should be cited and shouldn't not, e.g. excel} \LJHnote{sometimes excel could be considered materiel \ldots}
\niemnote{Example of using Excel to simply store and plot data, vs.\ using it for statistical analysis}
\ssnote{from the data side, an example might be found in the National Climate Assessment (NASA, GCIS) where this is captured in PROV instead?}
\dmnote{ Mention that the principles are covering what's necessary, not necessarily what is sufficient for a given use case. So perhaps we should discuss things like configurations or bugs? }
\paragraph{examples of citation styles}  \scnote{At least four styles (AMS, APA, Chicago, MLA) for
software citation are illustrated with examples in~\cite{lipson2011cite}.}

\paragraph{unique identifiers}  \katznote{recognize that a UID points to a specific software, but multiple UIDs might point to the same software} \LJHnote{not sure that multiple UIDs would be a recommended best practice as  this could cause conflicts} \katznote{LJH: not recommended, but likely will happen anyhow} \katznote{maybe discuss how identifiers are created}

\paragraph{talk about access to software - maybe discuss location}
\LJHnote{recognize that some software is proprietary, not all is open source}

\subsection{Reproducibility}  \katznote{recognize that more is needed for reproducibility, such as configuration information and runtime environment} \LJHnote{concept of enhance metadata to support this goal?}

\subsection{Updates}
\dmnote{Any provisions for updating the principles?}
\LJHnote{may require feedback from "Implementation Group" and some time to put into practice to find gaps. Most of the time revisions are frequent in the beginning of the lifecycle to cover gaps and become less frequent.}

>>>>

The principles in this section are written fairly concisely, and discussed further later in this document (\S\ref{sec:discussion}). For example, in this section we do not define what software should be cited, but how it should be cited.  we talk about how we think such decisions might be made in the discussion section.

@@

\textbf{What software to cite}

In the principles, we did not define what software should be cited, but how it should be cited.
What software should be cited is the decision of the author(s) of the research work, in the context of community norms and practices, and in most scientific communities, these are currently in flux.
In general, we believe that any software on which a research product directly depends should be cited, if it significantly contributes to the research product.
Again, the specific decision of what is significant must be made by the author(s) of the product.
However, an illustrative example is the use of Microsoft Excel in research.
We suggest that if Excel is used to simply store and plot data, it does not need to be cited, but if it is used for statistical analysis, it should be.
\ssnote{from the data side, an example might be found in the National Climate Assessment (NASA, GCIS) where this is captured in PROV instead?}
\dmnote{ Mention that the principles are covering what's necessary, not necessarily what is sufficient for a given use case. So perhaps we should discuss things like configurations or bugs? }

\textbf{Derived software}

@@

\textbf{Citations in text}

In terms of how citations should be expressed as text, such
as in journals or other human readable material, there are currently at least four styles (AMS, APA, Chicago, MLA). These have been illustrated with examples by Lipson~\cite{lipson2011cite}.
This is another area that is in flux at this time.

\textbf{Unique identifiers}

A UID points to a specific software, but multiple UIDs might point to the same software.
\LJHnote{not sure that multiple UIDs would be a recommended best practice as  this could cause conflicts} \katznote{LJH: not recommended, but likely will happen anyhow} \katznote{maybe discuss how identifiers are created}

\textbf{talk about access to software - maybe discuss location}

\LJHnote{recognize that some software is proprietary, not all is open source}

\textbf{Reproducibility}

\katznote{recognize that more is needed for reproducibility, such as configuration information and runtime environment} \LJHnote{concept of enhance metadata to support this goal?}

\textbf{Updates}

\dmnote{Any provisions for updating the principles?}
\LJHnote{may require feedback from "Implementation Group" and some time to put into practice to find gaps. Most of the time revisions are frequent in the beginning of the lifecycle to cover gaps and become less frequent.}
